---
title: "Step 2.6 - Databases using R"
author: "Jiaying Wu"
date: "06/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
library(tidyverse)
library(DBI)

con <- dbConnect(RSQLite::SQLite(), "data/:flights:")
flights_db <- tbl(con, "flights")
```

### Work with tidyverse wrangling verbs to access database

```{r}
df <- flights_db %>%
  group_by(OP_UNIQUE_CARRIER) %>%
  tally(sort=TRUE) %>%
  collect()
df
```
### What’s the smallest, largest and median departure delay for the busiest airport?

```{r}
df <- flights_db %>%
  filter(ORIGIN == "ATL") %>%
  collect()

df %>% summarise(min(DEP_DELAY, na.rm=TRUE), 
                 max(DEP_DELAY, na.rm=TRUE), 
                 median(DEP_DELAY, na.rm=TRUE))
```
### Make a side-by-side boxplot of the delays for each carrier, at the busiest airport.

```{r}
library(forcats)
df %>%
  mutate(DAY_OF_WEEK = as.factor(DAY_OF_WEEK)) %>%
  mutate(DepDelay_log = log10(DEP_DELAY+27)) %>%
  mutate(OP_CARRIER = fct_reorder(as_factor(OP_CARRIER), DepDelay_log, .desc = TRUE, na.rm=TRUE)) %>%
  ggplot(aes(x=OP_CARRIER, y=DepDelay_log)) + 
  geom_boxplot() + xlab("carrier") + ylab("Departure delay (log)")
```

### How many records of the busiest airport have missing values for departure delay?

```{r}
df %>% count(is.na(DEP_DELAY)) 
```

### Is there a ghost flight in your month of data?

```{r}
df <- flights_db %>%
  select(TAIL_NUM, ORIGIN, DEST, FL_DATE, DEP_TIME) %>%
  filter(!is.na(DEP_TIME)) %>%
  filter(!is.na(DEP_TIME)) %>%
  mutate(DEP_TIME = as.numeric(DEP_TIME)) %>%
  arrange(TAIL_NUM, FL_DATE, DEP_TIME) %>%
  collect()

df %>% count(TAIL_NUM, sort=TRUE)

aircraft <- df %>%
  select(TAIL_NUM) %>%
  distinct()

for (i in 1:nrow(aircraft)) {
  d <- df %>% filter(TAIL_NUM == aircraft$TAIL_NUM[i]) 
  if (nrow(d) > 1) {
    for (j in 2:nrow(d)) {
      if (d$ORIGIN[j] != d$DEST[j-1]) {
        cat(d$TAIL_NUM[j], d$DEST[j-1], d$ORIGIN[j],  "\n")
      }
    }
  }
  cat(i, "\n")
}
```

### Add tables to an existing database

Download the file from: https://www.transtats.bts.gov/DL_SelectFields.asp?gnoyr_VQ=FLL&QO_fu146_anzr=N8vn6v10%20f722146%20gnoyr5

```{r}
airports <- read_csv("data/654698093_T_MASTER_CORD.csv") %>%
  select(-...29)

copy_to(dbConnect(RSQLite::SQLite(), "data/airports"), airports, "airports", 
        temporary = FALSE
)

dbListTables(con)
airports_db <- tbl(con, "airports")
```

### Plot the locations on a map

```{r}
airport_locations <- airports_db %>%
  filter(AIRPORT_IS_LATEST == 1, AIRPORT_IS_CLOSED != 1,
         AIRPORT_COUNTRY_CODE_ISO == "US") %>%
  select(AIRPORT, DISPLAY_AIRPORT_NAME, LONGITUDE, LATITUDE) %>%
  collect() 

library(ggmap)
usa_bbox <- c(-130, # min long
              20, # min lat
              -60, # max long
              50) # max lat
usa_map <- get_map(location = usa_bbox, source = "stamen")

library(ggthemes)
# Only continental locations will be considered
airport_locations <- airport_locations %>%
  filter(LONGITUDE>usa_bbox[1], LONGITUDE<usa_bbox[3], 
         LATITUDE>usa_bbox[2], LATITUDE<usa_bbox[4])

ggmap(usa_map) + 
  geom_point(data=airport_locations, aes(x=LONGITUDE, y=LATITUDE),
             colour="red", alpha=0.3, size=1) +
  theme_map()
```

### Plot all flights

```{r}
# take a day's worth of flights
feb1 <- flights_db  %>%
  filter(DAY_OF_MONTH == 1) %>%
  collect()

# Remove flights that have a missing origin or dest lat/long
# because they are outside continental US
feb1_flights <- feb1 %>% 
  left_join(airport_locations, by=c("ORIGIN" = "AIRPORT")) %>% 
  rename(Origin_lon = LONGITUDE, Origin_lat = LATITUDE, 
         Origin_name = DISPLAY_AIRPORT_NAME) %>%
  left_join(airport_locations, by=c("DEST" = "AIRPORT")) %>%
  rename(Dest_lon = LONGITUDE, Dest_lat = LATITUDE, 
         Dest_name = DISPLAY_AIRPORT_NAME) %>%
  filter(!is.na(Origin_lon)) %>%
  filter(!is.na(Dest_lon)) %>%
  filter(!is.na(DEP_TIME))

ggmap(usa_map) + 
  geom_segment(data=feb1_flights, 
           aes(x=Origin_lon, 
               xend=Dest_lon,
               y=Origin_lat,
               yend=Dest_lat), 
           colour="#9651A0", alpha=0.01) +
  geom_point(data=feb1_flights, aes(x=Origin_lon, Origin_lat),
             colour="#746FB2", alpha=0.2, size=0.03) +
  theme_map()
```

### Compare carrier flight patterns

```{r}
feb1_major_carriers <- feb1_flights %>%
  filter(OP_CARRIER %in% c("WN", "AA")) %>%
  mutate(OP_CARRIER == fct_drop(OP_CARRIER))

ggmap(usa_map) + 
  geom_segment(data=feb1_major_carriers, 
           aes(x=Origin_lon, 
               xend=Dest_lon,
               y=Origin_lat,
               yend=Dest_lat), 
           colour="#9651A0", alpha=0.02) +
  geom_point(data=feb1_major_carriers, aes(x=Origin_lon, Origin_lat),
             colour="#746FB2", alpha=0.2, size=0.03) +
  facet_wrap(~OP_CARRIER, ncol=1) +
  theme_map()
```

### Examine change in patterns over the course of a day

```{r}
# Time zone matching
library(lutz)
library(lubridate)
feb1_flights <- feb1_flights %>% 
  mutate(origin_tz = tz_lookup_coords(Origin_lat, 
                              Origin_lon, warn=FALSE),
         dest_tz = tz_lookup_coords(Dest_lat, 
                              Dest_lon, warn=FALSE)) %>%
  mutate(DateTime = paste0("2020-02-01 ", 
                           substr(DEP_TIME, 1, 2),":", 
                           substr(DEP_TIME, 3, 4), ":00"))

# Can't keep datetime with different time zones in the same variable
feb1_flights <- feb1_flights %>%
  mutate(
    DateTime_NY = force_tzs(ymd_hms(DateTime),
      tzones = origin_tz, tzone_out = "America/New_York"))

# Now break into four time categories
feb1_flights <- feb1_flights %>% 
  mutate(TimeOfDay = case_when(
    hour(DateTime_NY) < 7 ~ "early",
    between(hour(DateTime_NY), 7, 11) ~ "morning",
    between(hour(DateTime_NY), 12, 17) ~ "afternoon",
    hour(DateTime_NY) > 17 ~ "evening")
  ) %>%
  mutate(TimeOfDay = factor(TimeOfDay, levels=c("early", "morning", "afternoon", "evening")))

# Track the path of one plane
ggmap(usa_map) + 
  geom_point(data=feb1_flights, aes(x=Origin_lon, y=Origin_lat),
             colour="red", alpha=0.2, size=0.5) +
  facet_wrap(~TimeOfDay, ncol=2) +
  theme_map()
```

### Use the standardised times to follow the path of one plane during the day

```{r}
N521AE <- feb1_flights %>% 
  filter(TAIL_NUM == "N521AE") %>%
  arrange(DateTime_NY)

ggmap(usa_map) + 
  geom_segment(data=N521AE, aes(x=Origin_lon, y=Origin_lat, 
                                xend=Dest_lon, yend=Dest_lat),
               colour="#9651A0", size = 1) +
  theme_map()
```






